Generalized Bradley-Terry models

### 2025-02-18 inst/jss_hyper3_real.Rnw
further work on probability tree diagram

\begin{figure}[htbp]
  \begin{center}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw (0,0) -- (2, 3) -- (4.8, 3);
\draw (0,0) -- (2, 2) -- (4.8, 2);
\draw (0,0) -- (2, 1) -- (4.8, 1);
\draw (0,0) -- (2, 0) -- (4.8, 0);
\draw (0,0) -- (2,-1) -- (4.8,-1);
\node at (0,2.5) (eq1) {$\underbrace{\left\lbrace a,b,c\right\rbrace}_{\lambda}\underbrace{\left\lbrace d,e\right\rbrace}_{\mu}$};

@@

\fill[red] (4.8, 3) circle[radius=2pt];  % a finishes; paths bcde
\fill[red] (4.8, 2) circle[radius=2pt];  % terminal node
\fill[red] (4.8, 1) circle[radius=2pt];  % terminal node
\fill[red] (4.8, 0) circle[radius=2pt];  % d finishes; paths abce
\fill[red] (4.8,-1) circle[radius=2pt];  % terminal node

\draw (4.8,3) -- (5.8,5) -- (8.2,5);
\draw (4.8,3) -- (5.8,4) -- (8.2,4);
\draw (4.8,3) -- (5.8,3) -- (8.2,3);
\draw (4.8,3) -- (5.8,2) -- (8.2,2);

@@

\fill[blue] (8.2,5) circle[radius=2pt];  % ab finishes; paths cde
\fill[blue] (8.2,4) circle[radius=2pt];  % ab finishes; paths cde
\fill[blue] (8.2,3) circle[radius=2pt];  % ab finishes; paths cde
\fill[blue] (8.2,2) circle[radius=2pt];  % ab finishes; paths cde
\draw (8.2,5) -- (8.9,6) -- (10.6,6);
\draw (8.2,5) -- (8.9,5) -- (10.6,5);
\draw (8.2,5) -- (8.9,4) -- (10.6,4);

@@

\fill[green] (10.6, 6) circle[radius=2pt];  % abc finishes; paths de
\draw (10.6, 6) -- (11.2,7) -- (12.2,7);
\fill[red] (12.2,7) circle[radius=2pt];
\node at (11.8, 7.4) {$\frac{p_d}{p_d+p_e}$};

\draw (10.6, 6) -- (11.2,6.2) -- (12.2,6.2);
\fill[red] (12.2,6.2) circle[radius=2pt];
\node at (11.8, 6.6) {$\frac{p_e}{p_d+p_e}$};

\fill[green] (10.6, 5) circle[radius=2pt];
\draw (10.6, 5) -- (11.2,5.3) -- (12.2, 5.3);
\fill[red] (12.2,5.3) circle[radius=2pt];
\node at (11.8, 5.7) {$\frac{\la{p_c}}{\la{p_c}+\ld{p_e}}$};

\draw (10.6, 5) -- (11.2,4.7) -- (12.2, 4.7);
\fill[red] (12.2,4.7) circle[radius=2pt];
\node at (11.8, 5.0) {$\frac{\ld{p_e}}{\la{p_c}+\ld{p_e}}$};

\fill[green] (10.6, 4) circle[radius=2pt];
\draw (10.6, 4) -- (11.2,3.9) -- (12.2, 3.9);
\fill[red] (12.2,3.9) circle[radius=2pt];
\node at (11.8, 4.3) {$\frac{\la{p_c}}{\la{p_c}+\ld{p_e}}$};

\draw (10.6, 4) -- (11.2,3.1) -- (12.2,3.1);
\fill[red] (12.2,3.1) circle[radius=2pt];
\node at (11.8, 3.5) {$\frac{\ld{p_e}}{\la{p_c}+\ld{p_e}}$};

\draw (4.8, 0) -- (5.8,1) -- (8.2,1);
\fill[red] (12.2,1) circle[radius=2pt];

\node at (7, 1.4)  {$\frac{p_a}{p_a+p_b+p_c+\ld{p_e}}$};

\draw (4.8, 0) -- (5.8,0) -- (8.2,0);
\fill[red] (12.2,0) circle[radius=2pt];
\node at (7, 0.4)  {$\frac{p_b}{p_a+p_b+p_c+\ld{p_e}}$};

\draw (4.8, 0) -- (5.8,-1) -- (8.2,-1);
\node at (7, -0.6)  {$\frac{p_c}{p_a+p_b+p_c+\ld{p_e}}$};

\draw (4.8, 0) -- (5.8,-2) -- (8.2,-2);
\node at (7, -1.6)  {$\frac{\ld{p_e}}{p_a+p_b+p_c+\ld{p_e}}$};

\fill[blue] (8.2, -2) circle[radius=2pt];  % de finishes; paths abc
\fill[blue] (8.2, -1) circle[radius=2pt];  % de finishes; paths abc
\fill[blue] (8.2,  0) circle[radius=2pt];  % de finishes; paths abc
\draw (8.2, -2) -- (8.9,-1) -- (10.6,-1);
\draw (8.2, -2) -- (8.9,-2) -- (10.6,-2);
\draw (8.2, -2) -- (8.9,-3) -- (10.6,-3);

\fill[red] (12.2,-1) circle[radius=2pt];

@@

\fill[green] (10.6, -1) circle[radius=2pt];  % dea finishes; paths bc
\fill[green] (10.6, -2) circle[radius=2pt];  % deb finishes; paths ac
\fill[green] (10.6, -3) circle[radius=2pt];  % dec finishes; paths ab

\draw (10.6, -1) -- (11.2,-0) -- (12.2,0);

\node at (11.8, 0.4) {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}}$};
\draw (10.6, -1) -- (11.2,-1) -- (12.2,-1);

@@

\draw(10.6, -2) -- (11.2,-1.8) -- (12.2,-1.8);
\fill[red] (12.2,-1.8) circle[radius=2pt];

\node at (11.8, -1.4) {$\frac{\la{p_a}}{\la{p_a}+\la{p_c}}$};

\draw (10.6, -2) -- (11.2,-2.6) -- (12.2,-2.6);
\fill[red] (12.2,-2.6) circle[radius=2pt];
\node at (11.8, -2.2) {$\frac{\la{p_c}}{\la{p_a}+\la{p_c}}$};

%\draw (10.6, -3) -- (11.2,-2.9);
%\fill[red] (12.2,-2.9) circle[radius=2pt];
%\node[red] at (11.8, -2.9) {$\frac{\la{p_a}}{\la{p_a}+\la{p_b}}$};

\draw (10.6, -3) -- (11.2,-3.4) -- (12.2, -3.4);
\fill[red] (12.2,-3.4) circle[radius=2pt];
\node at (11.8, -3) {$\frac{\la{p_b}}{\la{p_a}+\la{p_b}}$};

\fill[blue] (8.2, 1) circle[radius=2pt];  % da finishes; paths bce

\draw (8.2,1) -- (8.9,2) -- (10.6,2);
\draw (8.2,1) -- (8.9,1) -- (10.6,1);
\draw (8.2,1) -- (8.9,0) -- (10.6,0);

\node at (9.5, 2.4) {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}+p_e}$};
\node at (9.5, 1.4) {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}+p_e}$};
\node at (9.5, 0.4) {$\frac{p_e     }{\la{p_b}+\la{p_c}+p_e}$};
\fill[green] (10.6, 2) circle[radius=2pt];  % da finishes; paths bce
\fill[green] (10.6, 1) circle[radius=2pt];  % da finishes; paths bce
\fill[green] (10.6, 0) circle[radius=2pt];  % da finishes; paths bce
\draw (10.6, 1) -- (11.2,2) -- (12.2,2);
\node at (11.8, 2.4) {$\frac{\la{p_b}}{\la{p_b}+\ld{p_e}}$};
\draw (10.6, 1) -- (11.2,1) -- (12.2,1);
\node at (11.8, 1.4) {$\frac{\ld{p_e}}{\la{p_b}+p_e}$};

>>>>

\begin{figure}[htbp]
  \begin{center}
\usetikzlibrary{arrows}
\usetikzlibrary{patterns}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm]
\fill (0,0) circle[radius=2pt]; % root; paths abcde
\draw[thick] (0,0) -- (2, 3) -- (4.8, 3);
\draw[thick] (0,0) -- (2, 2) -- (4.8, 2);
\draw[thick] (0,0) -- (2, 1) -- (4.8, 1);
\draw[thick] (0,0) -- (2, 0) -- (4.8, 0);
\draw[thick] (0,0) -- (2,-1) -- (4.8,-1);
\node at (0,2.5) (eq1) {$\underbrace{\left\lbrace a,b,c\right\rbrace}_{\lambda}\underbrace{\left\lbrace d,e\right\rbrace}_{\mu}$};

@@

\fill (4.8, 3) circle[radius=2pt];  % a finishes; paths bcde
\fill (4.8, 2) circle[radius=2pt];  % terminal node
\fill (4.8, 1) circle[radius=2pt];  % terminal node
\fill (4.8, 0) circle[radius=2pt];  % d finishes; paths abce
\fill (4.8,-1) circle[radius=2pt];  % terminal node

\draw[thick] (4.8,3) -- (5.8,5) -- (8.2,5);
\draw[thick] (4.8,3) -- (5.8,4) -- (8.2,4);
\draw[thick] (4.8,3) -- (5.8,3) -- (8.2,3);
\draw[thick] (4.8,3) -- (5.8,2) -- (8.2,2);

@@

\fill (8.2,5) circle[radius=2pt];  % ab finishes; paths cde
\fill (8.2,4) circle[radius=2pt];  % ab finishes; paths cde
\fill (8.2,3) circle[radius=2pt];  % ab finishes; paths cde
\fill (8.2,2) circle[radius=2pt];  % ab finishes; paths cde

\node at (8.2,4.8) {$b$};
\node at (8.2,3.8) {$c$};
\node at (8.2,2.8) {$d$};
\node at (8.2,1.8) {$e$};

\draw[thick] (8.2,5) -- (8.9,6) -- (10.6,6);
\draw[thick] (8.2,5) -- (8.9,5) -- (10.6,5);
\draw[thick] (8.2,5) -- (8.9,4) -- (10.6,4);

@@

\fill (10.6, 6) circle[radius=2pt];  % abc finishes; paths de
\node at (10.6,5.8) {$c$};
\draw[thick] (10.6, 6) -- (11.2,7) -- (12.2,7);
\fill (12.2,7) circle[radius=2pt];
\node[anchor = west] at (12.2,7) {$a\succ b\succ c\succ d\succ e$};
\node at (11.8, 7.4) {$\frac{p_d}{p_d+p_e}$};

\draw[thick] (10.6, 6) -- (11.2,6.2) -- (12.2,6.2);
\fill (12.2,6.2) circle[radius=2pt];
\node[anchor = west] at (12.2,6.2) {$a\succ b\succ c\succ e\succ d$};
\node at (11.8, 6.6) {$\frac{p_e}{p_d+p_e}$};

\fill (10.6, 5) circle[radius=2pt];
\node at (10.6,4.8) {$d$};

\draw[thick] (10.6, 5) -- (11.2,5.3) -- (12.2, 5.3);
\fill (12.2,5.3) circle[radius=2pt];
\node[anchor = west] at (12.2,5.3) {$a\succ b\succ d\succ c\succ e$};
\node at (11.8, 5.7) {$\frac{\la{p_c}}{\la{p_c}+\ld{p_e}}$};

\draw[thick] (10.6, 5) -- (11.2,4.7) -- (12.2, 4.7);
\fill (12.2,4.7) circle[radius=2pt];
\node[anchor = west] at (12.2,4.7) {$a\succ b\succ d\succ e\succ c$};
\node at (11.8, 5.0) {$\frac{\ld{p_e}}{\la{p_c}+\ld{p_e}}$};

\fill (10.6, 4) circle[radius=2pt];
\node at (10.6,3.8) {$e$};
\draw[thick] (10.6, 4) -- (11.2,3.9) -- (12.2, 3.9);
\fill (12.2,3.9) circle[radius=2pt];
\node[anchor = west] at (12.2,3.9) {$a\succ b\succ e\succ c\succ d$};
\node at (11.8, 4.3) {$\frac{\la{p_c}}{\la{p_c}+\ld{p_d}}$};

\draw[thick] (10.6, 4) -- (11.2,3.1) -- (12.2,3.1);
\fill (12.2,3.1) circle[radius=2pt];
\node at (11.8, 3.5) {$\frac{\ld{p_d}}{\la{p_c}+\ld{p_d}}$};
\node[anchor = west] at (12.2,3.1) {$a\succ b\succ e\succ d\succ c$};

\draw[thick] (4.8, 0) -- (5.8,1) -- (8.2,1);
\fill (12.2,1) circle[radius=2pt];
\node[anchor = west] at (12.2,1) {$d\succ a\succ c\succ e\succ b$};
\node at (7, 1.4)  {$\frac{p_a}{p_a+p_b+p_c+\ld{p_e}}$};

\draw[thick] (4.8, 0) -- (5.8,0) -- (8.2,0);
\fill (12.2,0) circle[radius=2pt];
\node[anchor = west] at (12.2,0) {$d\succ e\succ a\succ b\succ c$};
\node at (7, 0.4)  {$\frac{p_b}{p_a+p_b+p_c+\ld{p_e}}$};

\draw[thick] (4.8, 0) -- (5.8,-1) -- (8.2,-1);
\node at (7, -0.6)  {$\frac{p_c}{p_a+p_b+p_c+\ld{p_e}}$};

\draw[thick] (4.8, 0) -- (5.8,-2) -- (8.2,-2);
\node at (7, -1.6)  {$\frac{\ld{p_e}}{p_a+p_b+p_c+\ld{p_e}}$};

\fill (8.2, -2) circle[radius=2pt];  % de finishes; paths abc
\fill (8.2, -1) circle[radius=2pt];  % de finishes; paths abc
\fill (8.2,  0) circle[radius=2pt];  % de finishes; paths abc

\node at (8.2, -2.2) {$e$};
\node at (8.2, -1.2) {$c$};
\node at (8.2, -0.2) {$b$};
\node at (8.2,  0.8) {$a$};

\draw[thick] (8.2, -2) -- (8.9,-1) -- (10.6,-1);
\draw[thick] (8.2, -2) -- (8.9,-2) -- (10.6,-2);
\draw[thick] (8.2, -2) -- (8.9,-3) -- (10.6,-3);

\fill (12.2,-1) circle[radius=2pt];
\node[anchor = west] at (12.2,-1) {$d\succ e\succ a\succ c\succ b$};

@@

\fill (10.6, -1) circle[radius=2pt];  % dea finishes; paths bc
\fill (10.6, -2) circle[radius=2pt];  % deb finishes; paths ac
\fill (10.6, -3) circle[radius=2pt];  % dec finishes; paths ab
\node at (10.6,-1.2) {$a$};
\node at (10.6,-2.2) {$b$};
\node at (10.6,-3.2) {$c$};

\draw[thick] (10.6, -1) -- (11.2,-0) -- (12.2,0);

\node at (11.8, 0.4) {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}}$};
\draw[thick] (10.6, -1) -- (11.2,-1) -- (12.2,-1);

@@

\draw[thick](10.6, -2) -- (11.2,-1.8) -- (12.2,-1.8);
\fill (12.2,-1.8) circle[radius=2pt];
\node[anchor = west] at (12.2,-1.8) {$d\succ e\succ b\succ a\succ c$};
\node at (11.8, -1.4) {$\frac{\la{p_a}}{\la{p_a}+\la{p_c}}$};

\draw[thick] (10.6, -2) -- (11.2,-2.6) -- (12.2,-2.6);
\fill (12.2,-2.6) circle[radius=2pt];
\node[anchor = west] at (12.2,-2.6) {$d\succ e\succ b\succ c\succ a$};
\node at (11.8, -2.2) {$\frac{\la{p_c}}{\la{p_a}+\la{p_c}}$};

%\draw[thick] (10.6, -3) -- (11.2,-2.9);
%\fill (12.2,-2.9) circle[radius=2pt];
%\node at (11.8, -2.9) {$\frac{\la{p_a}}{\la{p_a}+\la{p_b}}$};

\draw[thick] (10.6, -3) -- (11.2,-3.4) -- (12.2, -3.4);
\fill (12.2,-3.4) circle[radius=2pt];
\node[anchor = west] at (12.2,-3.4) {$d\succ e\succ c\succ b\succ a$};

\node at (11.8, -3) {$\frac{\la{p_b}}{\la{p_a}+\la{p_b}}$};

\fill (8.2, 1) circle[radius=2pt];  % da finishes; paths bce

\draw[thick] (8.2,1) -- (8.9,2) -- (10.6,2);
\draw[thick] (8.2,1) -- (8.9,1) -- (10.6,1);
\draw[thick] (8.2,1) -- (8.9,0) -- (10.6,0);

\node at (9.5, 2.4) {$\frac{\la{p_b}}{\la{p_b}+\la{p_c}+p_e}$};
\node at (9.5, 1.4) {$\frac{\la{p_c}}{\la{p_b}+\la{p_c}+p_e}$};
\node at (9.5, 0.4) {$\frac{p_e     }{\la{p_b}+\la{p_c}+p_e}$};
\fill (10.6, 2) circle[radius=2pt];  % da finishes; paths bce
\fill (10.6, 1) circle[radius=2pt];  % da finishes; paths bce
\fill (10.6, 0) circle[radius=2pt];  % da finishes; paths bce

\node at (10.6, 1.8) {$b$};
\node at (10.6, 0.8) {$c$};
\node at (10.6,-0.2) {$e$};

\draw[thick] (10.6, 1) -- (11.2,2) -- (12.2,2);
\node at (11.8, 2.4) {$\frac{\la{p_b}}{\la{p_b}+\ld{p_e}}$};
\fill (12.2,2) circle[radius=2pt];
\node[anchor = west] at (12.2,2) {$d\succ a\succ c\succ b\succ e$};
\draw[thick] (10.6, 1) -- (11.2,1) -- (12.2,1);
\node at (11.8, 1.4) {$\frac{\ld{p_e}}{\la{p_b}+p_e}$};
